Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-08-03 Manuscript/journal_correspondence/ERL_responseletter.Rmd

---
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
header-includes:
   - \linespread{1.05}
   - \newcommand{\blankline}{\quad\pagebreak[2]}
   - \usepackage{eso-pic}
   - \usepackage{graphicx}
   - \newcommand\BackgroundPic{\put(0,0){\parbox[b][\paperheight]{\paperwidth}{\vfill\centering \includegraphics[width=\paperwidth,height=\paperheight, keepaspectratio]{usgs_response_letterhead.pdf} \vfill }}}
geometry: top=1in, left=1in, bottom=1in, right=1in
mainfont: Times New Roman
fontsize: 12pt
urlcolor: blue
---

\AddToShipoutPicture*{\BackgroundPic}
\pagenumbering{gobble}

`r format(Sys.time(), '%B %d, %Y')`

\blankline

Dear Editors:

\blankline

Please consider our article, "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of climate and land use change", for publication in _Environmental Research Letters_. This article presents results from a stochastic, spatially explicit simulation model used to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands over a 90-year timespan from 2010-2100. Our results represent the first set of Hawaiʻi-specific projections to incorporate multiple land use and climate change scenarios, the first to  extend to the end of the 21st century, and the first to be conducted at sub-kilometer spatial resolution. This article also represents one of very few regional-scale efforts to examine how ecosystem carbon balance responds to different levels of a CO~2~ fertilization effect on net primary productivity.

Our results are timely given the recent passage of legislation by the State of Hawaiʻi to be carbon neutral by 2045, a goal that will partly depend on carbon sequestration by terrestrial ecosystems. By incorporating the interactive effects of land use and climate change into future projections of ecosystem carbon balance, our model results could serve as baseline projections for the State of Hawaiʻi to evaluate different ecosystem-based climate mitigation strategies. Studies like ours that incorporate stochasticity into spatially explicit simulation models could also provide a framework for the growing number of sub-national jurisdictions that plan to incorporate ecosystem carbon sequestration into their emissions reduction efforts.

We found that local reductions in the intensity of land use change consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global climate mitigation through greenhouse gas emissions reductions. We also found that CO~2~ fertilization of NPP was the largest source of uncertainty in long-term projections of ecosystem carbon balance in the Hawaiian Islands, highlighting the need for greater mechanistic understanding of the cascading effects of rising atmospheric CO~2~ on ecosystem carbon sequestration. We believe long-term model projections such as ours will be critical to understanding how future land use and climate change could interact to influence the achievement of climate mitigation goals.

\blankline

Best regards,

\raggedright
Paul C. Selmants 
\linebreak
Research Ecologist 
\linebreak
U.S. Geological Survey

Lots and lots more text, repeating last paragraph of letter here to see what happens on page two.

We found that local reductions in the intensity of land use change consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global climate mitigation through greenhouse gas emissions reductions. We also found that CO~2~ fertilization of NPP was the largest source of uncertainty in long-term projections of ecosystem carbon balance in the Hawaiian Islands, highlighting the need for greater mechanistic understanding of the cascading effects of rising atmospheric CO~2~ on ecosystem carbon sequestration. We believe long-term model projections such as ours will be critical to understanding how future land use and climate change could interact to influence the achievement of climate mitigation goals.

### 2021-08-03 Manuscript/HI_Model_ms.Rmd
updated journal correspondence directory

We used a stochastic, spatially explicit simulation model to estimate ecosystem carbon balance for Hawaiʻi's natural and agricultural lands on an annual basis for the period 2010–2100 under a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions [@daniel_state-and-transition_2016; @daniel_integrating_2018; @sleeter_effects_2019]. We explored four unique scenarios that represent all combinations of two land use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and RCP 8.5). In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019]. Our goals were to estimate changes in Hawaiian ecosystem carbon balance and their uncertainties under a range of plausible future scenarios, quantify the relative impact of major controlling processes such as land use change, disturbance, and climate change, and assess the sensitivity of model estimates to the introduction of a CFE on NPP.

## 2. Methods
We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands. The landscape change portion of LUCAS is a state-and-transition model, where 'state' is the pre-defined land cover class  assigned to each simulation cell and 'transition' is the change in state of each simulation cell over time [@daniel_state-and-transition_2016]. Landscape change occurs by applying a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transition pathways [@daniel_state-and-transition_2016]. The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying rates of change in stock levels over time [@daniel_integrating_2018; @sleeter_effects_2019]. We parameterized the Hawaiʻi LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability for the time period 2010-2100.

@@

### *2.2 States and transitions*
We developed two land use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016. Transition pathways between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire. Agriculture, forest, grassland, shrubland, and tree plantation state types each had multiple transition pathways, while the barren state type could only transition to developed (i.e., urbanization). Although most state types had an urbanization transition pathway, there was no transition pathway out of an urbanized (developed) state. Water and wetland state types remained static throughout the simulation period.

 Transition targets set the area to be transitioned over time. These targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018]. For the high land use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island. Urbanization rates in the low land use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land use scenario and ~40% of recent historical rates in the low land use scenario [@daniel_state-and-transition_2016]. In both land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture. The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.

>>>>

We used a stochastic, spatially explicit simulation model to estimate ecosystem carbon balance for Hawaiʻi's natural and agricultural lands on an annual basis for the period 2010–2100 under a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions [@daniel_state-and-transition_2016; @daniel_integrating_2018; @sleeter_effects_2019]. We developed four unique scenarios that explored different pathways, or future trajectories, of land use change and climate change. These four scenarios represent all combinations of two land use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and RCP 8.5). In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019]. Our goals were to estimate changes in Hawaiian ecosystem carbon balance and their uncertainties under a range of plausible future scenarios, quantify the relative impact of major controlling processes such as land use change, disturbance, and climate change, and assess the sensitivity of model estimates to the introduction of a CFE on NPP.

## 2. Methods
We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands. The landscape change portion of LUCAS is a state-and-transition model, where 'state' is the pre-defined information specific to each simulation cell (e.g., land cover class, climate zone) and 'transition' is the change in state of each simulation cell over time [@daniel_state-and-transition_2016]. Landscape change occurs by applying a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transition pathways [@daniel_state-and-transition_2016]. The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying rates of change in stock levels over time [@daniel_integrating_2018; @sleeter_effects_2019]. We parameterized the Hawaiʻi LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability for the time period 2010-2100.

@@

### *2.2 States and transitions*
We developed two land use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016. Transition pathways define which state types can be converted to any other state type. Transition pathways between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire. Agriculture, forest, grassland, shrubland, and tree plantation state types each had multiple transition pathways, i.e., they could each transition to multiple other state types. The barren state type could only transition to developed, so its sole transition pathway was urbanization. Although most state types had an urbanization transition pathway, there was no transition pathway out of an urbanized (developed) state. Water and wetland state types remained static throughout the simulation period.

 Transition targets set the area to be transitioned over time. These targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018]. For the high land use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island. Urbanization rates in the low land use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land use scenario and ~40% of recent historical rates in the low land use scenario [@daniel_state-and-transition_2016]. In both land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture. The remaining 40% were rotation harvests replanted to *Eucalyptus* spp., which represents the vast majority of plantation forestry in the State of Hawaiʻi.
